---
title: "Netcdf Data Processing"
output: html_notebook
---

This notebook process raw Netcdf codes from Randall Martin's surface PM2.5 exposure dataset(https://sites.wustl.edu/acag/datasets/surface-pm2-5/)

```{r}
library(ncdf4)
#library(raster)
library(tidyverse)
#library(geometry)
#library(data.table)
#library(maps)
library(magrittr)
library(ggplot2)
library(sf)
library(reshape2) # For reshaping data
#library(mapdata)  # For map data
#library(ggmap)
library(lubridate)
library(raster)
#library(rgeos)
library(viridis)
library(tictoc)
```

```{r}
grid_sf <- read_sf("../data/input/10km_grid_wgs84/10km_grid_wgs84.shp")

st_crs(grid_sf)

grid_sf[] %>% 
  ggplot() + 
  geom_sf(fill = "blue", lwd = 0.00)
```


```{r}
nc_file <- nc_open("../data/input/pm25_components/PM25/V4NA03_PM25_NA_200001_200012-RH35.nc")
print(nc_file)
```

```{r}
attributes(nc_file)
```

```{r}
pm25 <- ncvar_get(nc_file, "PM25")
class(pm25)
dim(pm25)
dim(pm25)[1]*dim(pm25)[2]
sum(!is.na(pm25))
sum(is.na(pm25))
```

```{r}
pm25[1:10, 1:10]
```

```{r}
lon <- ncvar_get(nc_file, "LON")
lat <- ncvar_get(nc_file, "LAT")
class(lon)
length(lon)
class(lat)
length(lat)

lon[1:10]
lat[1:10]
```

```{r}
# Reshape data
pm25_df <- melt(pm25)
pm25_df
```

```{r}
pm25_df$lon <- rep(lon, length(lat))
pm25_df$lat <- rep(lat, each = length(lon))
pm25_df
```

```{r}
pm25_df <- pm25_df %>% 
  dplyr::select(-c(Var1, Var2)) %>% 
  dplyr::rename(pm25 = value) %>% 
  drop_na(pm25)

dim(pm25_df)
```

```{r}
pm25_df
```

```{r}
pm25_sf <- st_as_sf(pm25_df, coords = c("lon","lat"))
head(pm25_sf)
```

```{r}
st_crs(pm25_sf) <- st_crs(grid_sf)
head(pm25_sf)
```

```{r}
grid_sf %>% 
  ggplot() + 
  geom_sf(fill = "blue", lwd = 0.00) + 
  geom_sf(data = sample_n(pm25_sf, 10000), 
          color = "red", 
          size = 0.5)
```


```{r}
sample_pm25 <- sample_n(pm25_sf, 10000)

grid_sf %>% 
  ggplot() +
  geom_sf(fill = "pink", lwd = 0.00) + 
  geom_sf(data = sample_pm25, aes(color = pm25), size = 0.5) +
  scale_color_distiller(palette = "Spectral")
```

```{r}
rm(pm25)
rm(pm25_df)
gc()
```


```{r}
#sample_pm25 <-  sample_n(pm25_sf, 1000000)
```

```{r}
grid_i_sf <- dplyr::select(grid_sf[1:100, ], c("geometry"))
#dots_inside <- st_contains(first_zip ,, sparse = TRUE)
pm25_i_sf <- dplyr::select(st_crop(pm25, extent(grid_i_sf)), c("pm25", "geometry"))
```

```{r}
grid_i_sf %>% 
  ggplot() + 
  geom_sf() + 
  geom_sf(data = pm25_i_sf, color = "red")
```


```{r}
pm25_i_sp = as_Spatial(pm25_i_sf)

for (j in 1:100) {
  grid_j_sf <- dplyr::select(grid_i_sf[j, ], c("geometry"))

  if(class(grid_j_sf)[1] != "try-error"){
    grid_j_sp = as_Spatial(grid_j_sf)

    pm25 <- as.numeric(grid_j_sp %over% pm25_i_sp)
  } else pm25 <- as.numeric(NA)
  
  print(pm25)
}

```



# 1. select points that are in the USA 
# 2. Over function (faster than interpolate ) from 1 km points —> grid values for 10km grid squares 
Get 10km grid squares from Randall Martin’s code 
# 3. Over —> take value to 1x1 km grid —> convert to 10x10 km grid —> use smoke_aggregation workflow 
```{r}
grid_sf[1:100,] %>% 
  ggplot() +
  geom_sf()
```
```{r}
zip_sf = read_sf("../data/input/zipcode/polygon/ESRI06USZIP5_POLY_WGS84.shp")
st_crs(zip_sf)
st_crs(grid_sf)
```

```{r}
first_zip <- zip_sf[5000:6000,]
dots_inside <- st_contains(first_zip ,sample_pm25, sparse = TRUE)
inside_dots <- which(dots_inside)
inside_dots

first_zip%>% 
  ggplot() + 
  geom_sf(aes(fill = "red"), alpha = 0.75, lwd = 0.1) + 
  theme(legend.position = "none")
```
```{r}

insides <- sample_pm25[dots_inside,]
# Plot the points inside the first zip code area
ggplot() + 
  geom_sf(data = first_zip, fill = "transparent", color = "red", size = 1) +
  geom_sf(data = dots_inside, color = "blue", size = 1) +
  coord_sf() +
  labs(title = "Points Inside First Zip Code Area")
```



```{r}
zip_sf %>% 
  ggplot() + 
  geom_sf(aes(fill = "red"), alpha = 0.75, lwd = 0.1) + 
  theme(legend.position = "none")
```
```{r}
ext.ras <- extent(grid_sf)
ext.pol <- extent(zip_sf)

plot(ext.ras, 
     xlim = c(min(ext.ras@xmin, ext.pol@xmin), 
              max(ext.ras@xmax, ext.pol@xmax)), 
     ylim= c(min(ext.ras@ymin, ext.pol@ymin), 
              max(ext.ras@ymax, ext.pol@ymax)), 
     col="red")
plot(ext.pol, add=T, col="blue")
```
```{r}
# 3245 causes the error 
zip_sf[3245,]
sf_use_s2(FALSE)
zip_sf2 <- st_crop(zip_sf[3250:10000,], st_bbox(grid_sf))

zip_sf2 %>% 
  ggplot() + 
  geom_sf(aes(fill = "red"), alpha = 0.75, lwd = 0.1) + 
  theme(legend.position = "none")

```
```{r}
zip_sf2 <- data.frame()

error_indexes <- c()

for (i in 1:5000) {
  tryCatch({
    crop_zip_sf <- st_crop(zip_sf[i,], st_bbox(grid_sf))
    zip_sf2 <- rbind(zip_sf2, crop_zip_sf)
  }, error = function(e) {
    error_indexes <- c(error_indexes, i)
  })
}

```

```{r}
# determine which points from sample_pm25 fall within the bounding box of grid_sf
pm25_within_bbox <- sample_pm25[st_intersects(sample_pm25, st_bbox(grid_sf)), ]
```


For grid I, select geometry, from all the points, we identify which points 

Take a grid and crop all the points that fall within the bounding box of a give n single grid cells 

Cut all the points falling into bounding box (extent) —> apply over to give single value for that grid —> assign average value to that grid 

```{r}
head(sample_pm25)
# Define the grid cell size in meters (10 km x 10 km)
cell_size <- 10000

# Group the points into grid cells and calculate the mean value of pm25 for each cell
sample_pm25_grid <- sample_pm25 %>%
  st_transform(crs = st_crs("+proj=utm +zone=10 +datum=WGS84")) %>% # convert to UTM for distance calculation
  st_make_grid(cell_size = cell_size) %>%
  st_intersection(sample_pm25) %>%
  group_by(id) %>%
  summarise(pm25_mean = mean(pm25, na.rm = TRUE))

# Print the resulting data frame
print(sample_pm25_grid)
```
```{r}
library(sf)

# Convert sample_pm25 to an sf object
sample_pm25_sf <- st_as_sf(sample_pm25, coords = c("lon", "lat"), crs = 4326)

# Group points into a 10x10 km grid
grid_size <- 10 # meters
sample_pm25_grid <- st_make_grid(sample_pm25_sf, cellsize = c(grid_size, grid_size))

# Assign points to grid cells
sample_pm25_grid$pm25 <- st_extract(sample_pm25_sf["pm25"], sample_pm25_grid)

# Calculate the mean pm25 value for each grid cell
sample_pm25_grid_mean <- aggregate(sample_pm25_grid["pm25"], sample_pm25_grid, mean)

# Print the resulting data frame
print(sample_pm25_grid_mean)

```

```{r}
sfc = st_sfc(st_polygon(list(rbind(c(0,0), c(1,0), c(1,1), c(0,0)))))
plot(sample_pm25_sf)
plot(sample_pm25_grid)
```
```{r}

# Group points into a 10x10 km grid
grid_size <- 100 # meters 

polygony <- st_make_grid(sample_pm25_sf, square = T, cellsize = c(grid_size, grid_size)) %>% # the grid, covering bounding box
  st_sf() # not really required, but makes the grid nicer to work with later

plot(polygony, col = 'white')
plot(st_geometry(sample_pm25_sf), add = T)

```

```{r}
grid_sf = read_sf("../data/input/remote_data/10km_grid_wgs84/10km_grid_wgs84.shp")



```

